---
title: "Visualizing DaRTS CTD data from a single cruise"
output: 
  html_notebook: default
  html_document: default
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Workthrough of the practice questions

## Loading your data into R

First we want to load the R libraries we'll be using today:

```{r}
library(tidyverse)
```

```{r}
data <- read.csv("../DaRTS-data/DaRTS_SingleCruise_CTDdata.csv", header = TRUE)
```


## 1. Plot salinity, density and pH profiles for each station using faceting

### salinity
```{r}
ggplot(data, aes(x = sal_PSU, y = depth_m)) +
  geom_point() + 
  facet_wrap(~station, ncol = 4) +
  scale_y_reverse()
```

### density
```{r}
ggplot(data, aes(x = density_kg_m.3, y = depth_m)) +
  geom_point() + 
  facet_wrap(~station, ncol = 4) +
  scale_y_reverse()
```
### pH

```{r}
ggplot(data, aes(x = pH, y = depth_m)) +
  geom_point() + 
  facet_wrap(~station, ncol = 4) +
  scale_y_reverse()
```

## 2. For each of the following variables, plot profiles for each station on one plot: PAR, fluorescence, turbidity and O2 percent saturation

### PAR
```{r}
ggplot(data, aes(x = PAR, y = depth_m, color = factor(station))) +
  geom_point(size = 1) + 
  scale_y_reverse() + 
  labs(color = 'Station')
```

### fluorescence
```{r}
ggplot(data, aes(x = fluor_mg_m.3, y = depth_m, color = factor(station))) +
  geom_point(size = 1) + 
  scale_y_reverse() + 
  labs(color = 'Station')
```

### turbidity
```{r}
ggplot(data, aes(x = turbidity_NTU, y = depth_m, color = factor(station))) +
  geom_point(size = 1) + 
  scale_y_reverse() + 
  labs(color = 'Station')
```

### o2 saturation
```{r}
ggplot(data, aes(x = O2sat_percent, y = depth_m, color = factor(station))) +
  geom_point(size = 1) + 
  scale_y_reverse() + 
  labs(color = 'Station')
```

## Bonus questions (Google is your friend): Can you change the color, shape and/or size of the dots in the facet plot?

For color options, type color() at the console.
```{r}
ggplot(data, aes(x = temp_C, y = depth_m)) +
  geom_point(colour="thistle") + 
  facet_wrap(~station, ncol = 4) +
  ylim(100, 0) + xlim(7.5, 18)
```

link for shapes: https://ggplot2.tidyverse.org/reference/aes_linetype_size_shape.html?q=shape%20#null

```{r}
ggplot(data, aes(x = temp_C, y = depth_m)) +
  geom_point(shape=6) + 
  facet_wrap(~station, ncol = 4) +
  ylim(100, 0) + xlim(7.5, 18)
```

```{r}
ggplot(data, aes(x = temp_C, y = depth_m)) +
  geom_point(shape=15, size=5, color="green") + 
  facet_wrap(~station, ncol = 4) +
  ylim(100, 0) + xlim(7.5, 18)
```
b. Can you change the color, shape and/or size of the dots in the single plot?



