---
title: "Untitled"
output: html_document
date: "2025-09-10"
---
```{r}
library(ggplot2)
library(dplyr)
library(lubridate)
```


# 2024 

```{r}
DATA2024 <- read.csv("../../2024_sea_change_rlab/Lab_5/DaRTS_combined_data.csv")
```

```{r}
DATA2024 <- DATA2024 %>% mutate(Date = lubridate::ymd(DATA2024$Date))
```

Our data frame contains data from 2012 - 2020, so we need to `filter` the data frame based on year and depth:

```{r}
chldata2016_2024 <- DATA2024 %>% filter(year == 2016, Depth == 1)
```


## Visualizing data using a contour plot

We can visualize this type of data with a contour plot (using `geom_contour_filled()`). We want to label the color bar, so in this case, our argument to the `labs()` function for the legend (color bar) is going to be `fill`.

```{r, fig.height = 5, fig.width=8}
ggplot(chldata2016_2024,aes(x=Date,y=Station)) +
  geom_contour_filled(aes(z=Fluorescence)) +
  geom_point() +
  labs(fill='surface chlorophyll fluorescence (mg m^-3)')
```

# 2025 

```{r}
DATA2025 <- read.csv("data/DaRTS_combined_data.csv")
```

```{r}
DATA2025 <- DATA2025 %>% mutate(Date = lubridate::ymd(DATA2025$Date))
```

Our data frame contains data from 2012 - 2020, so we need to `filter` the data frame based on year and depth:

```{r}
chldata2016_2025 <- DATA2025 %>% filter(year == 2016, Depth == 1)
```


```{r}
chldata2016_2025 <- chldata2016_2025 %>% mutate(Station = as.numeric(chldata2016_2025$Station))
```


## Visualizing data using a contour plot

We can visualize this type of data with a contour plot (using `geom_contour_filled()`). We want to label the color bar, so in this case, our argument to the `labs()` function for the legend (color bar) is going to be `fill`.

```{r, fig.height = 5, fig.width=8}
ggplot(chldata2016_2025,aes(x=Date,y=Station)) +
  geom_contour_filled(aes(z=Fluorescence)) +
  geom_point() +
  labs(fill='surface chlorophyll fluorescence (mg m^-3)')
```